---
title: "Advanced Regression Assignment 6"
author: "Cullen Blair"
date: "11/22/2020"
output: word_document
---
library(faraway)
oh.df=ohio
names(oh.df)
n=nrow(oh.df)
calc.out = function(e,s)
  round(cbind(est=e,se=s,lower=e-qnorm(0.975)*s,upper=e+qnorm(0.975)*s),5)
sum.o = function(fit)
  round(cbind(summary(fit)$coefficients,confint.default(fit)),digits=5)
phi.f = function(c,p) c(p=p,est=c/(n-p),c=c,p.value=1-pchisq(c,n-p))
oh.df$id=as.factor(oh.df$id)
y=oh.df$resp
#1a
table(oh.df$age,oh.df$smoke)
gs = aggregate(y,by=list(oh.df$age,oh.df$smoke),mean)
x.tab = xtabs(x~Group.1+Group.2,data=gs)
names(dimnames(x.tab)) = c("age","smoke")
x.tab
#1b
#fit.glm.bin = glm(y~age+smoke,data=oh.df,family=binomial)
fit.glm.bin = glm(resp~age+smoke+age*smoke,data=oh.df,family=binomial) #no evidence against null of 0 interaction
sum.o(fit.glm.bin)  # [ELM, p 287] #
c(aic=AIC(fit.glm.bin,k=2),bic=AIC(fit.glm.bin,k=log(n)))
pearson = sum(resid(fit.glm.bin,type="pearson")^2)
d.test = phi.f(pearson,length(fit.glm.bin$coefficients))
d.test

#2a
library(lme4)
sum.mer = function(mer) round(cbind(summary(mer)$coefficients,
                                    confint(mer,method="Wald")[-c(1:length(mer@theta)),]),digits=5)
fit.glmer = glmer(y~age+smoke+age*smoke+(1|id),data=oh.df,family=binomial)

sum.mer(fit.glmer) 
summary(fit.glmer)$AICtab
vc = as.numeric(VarCorr(fit.glmer)$id) # vc = 0.5149987
H = fit.glmer@optinfo$derivs$Hessian # hessian # 
se = sqrt(diag(2*solve(H))) # std errors #
calc.out(sqrt(vc),se[1])
#2b
library(msm) # deltamethod #
d.se = deltamethod(~x1^2,sqrt(vc),se[1]^2,ses=T)
calc.out(vc,d.se) # 2*sqrt(vc)*se[1] #
#2c
sum.mer(fit.glmer)
#3a
library(geepack)
gee.fit <- geeglm(y ~age+smoke+I(age*smoke),data=oh.df,family=binomial, id=id, corstr="ar1")
summary(gee.fit)
summary(gee.fit)$AICtab
calc.out(summary(gee.fit)$corr[1],summary(gee.fit)$corr[2])
#3b
library(lme4)
sum.mer = function(mer) round(cbind(summary(mer)$coefficients,
                                    confint(mer,method="Wald")[-c(1:length(mer@theta)),]),digits=5)
fit.glmer2 = glmer(resp~smoke+age+smoke*age+(age-1|id)+(1|id),data=oh.df,family=binomial)
# sfit.glmer = summary(fit.glmer) # [ELM, p 288] #
sum.mer(fit.glmer2) # nAGQ=25 #
summary(fit.glmer2)$AICtab

library(lmtest)
lrtest(fit.glmer,fit.glmer2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(faraway)
oh.df=ohio
names(oh.df)
n=nrow(oh.df)
calc.out = function(e,s)
  round(cbind(est=e,se=s,lower=e-qnorm(0.975)*s,upper=e+qnorm(0.975)*s),5)
sum.o = function(fit)
  round(cbind(summary(fit)$coefficients,confint.default(fit)),digits=5)
phi.f = function(c,p) c(p=p,est=c/(n-p),c=c,p.value=1-pchisq(c,n-p))
oh.df$id=as.factor(oh.df$id)
y=oh.df$resp
```

```{r}
table(oh.df$age,oh.df$smoke)
gs = aggregate(y,by=list(oh.df$age,oh.df$smoke),mean)
x.tab = xtabs(x~Group.1+Group.2,data=gs)
names(dimnames(x.tab)) = c("age","smoke")
x.tab
```

```{r}
fit.glm.bin = glm(resp~age+smoke+age*smoke,data=oh.df,family=binomial) #no evidence against null of 0 interaction
sum.o(fit.glm.bin)  # [ELM, p 287] #
c(aic=AIC(fit.glm.bin,k=2),bic=AIC(fit.glm.bin,k=log(n)))
pearson = sum(resid(fit.glm.bin,type="pearson")^2)
d.test = phi.f(pearson,length(fit.glm.bin$coefficients))
d.test
```


```{r}
library(lme4)
sum.mer = function(mer) round(cbind(summary(mer)$coefficients,
                                    confint(mer,method="Wald")[-c(1:length(mer@theta)),]),digits=5)
fit.glmer = glmer(y~age+smoke+age*smoke+(1|id),data=oh.df,family=binomial)

sum.mer(fit.glmer) 
summary(fit.glmer)$AICtab
vc = as.numeric(VarCorr(fit.glmer)$id) # vc = 0.5149987
H = fit.glmer@optinfo$derivs$Hessian # hessian # 
se = sqrt(diag(2*solve(H))) # std errors #
calc.out(sqrt(vc),se[1])
```

```{r}
library(msm) # deltamethod #
d.se = deltamethod(~x1^2,sqrt(vc),se[1]^2,ses=T)
calc.out(vc,d.se)
```

```{r}
sum.mer(fit.glmer)
```

```{r}
library(geepack)
gee.fit <- geeglm(y ~age+smoke+I(age*smoke),data=oh.df,family=binomial, id=id, corstr="ar1")
summary(gee.fit)
summary(gee.fit)$AICtab
calc.out(summary(gee.fit)$corr[1],summary(gee.fit)$corr[2])
```

```{r}
library(lme4)
sum.mer = function(mer) round(cbind(summary(mer)$coefficients,
                                    confint(mer,method="Wald")[-c(1:length(mer@theta)),]),digits=5)
fit.glmer2 = glmer(resp~smoke+age+smoke*age+(1|id)+(age-1|id),data=oh.df,family=binomial)
sum.mer(fit.glmer2)
vc2.0 = as.numeric(VarCorr(fit.glmer2)$id)
vc2.1= as.numeric(VarCorr(fit.glmer2)$id.1)
H2 = fit.glmer2@optinfo$derivs$Hessian # hessian # 
se2 = sqrt(diag(2*solve(H2))) # std errors #
calc.out(sqrt(vc2.0),se2[1])
calc.out(sqrt(vc2.1),se2[2])
summary(fit.glmer2)$AICtab
```
```{r}
anova(fit.glmer,fit.glmer2)
```
